A Fast kNN Algorithm Using Multiple Space-Filling Curves

### 2022-04-04 2022/Sensors/sensors.tex
Results and Conclusion

The study was performed using several datasets, both randomly generated for testing and real-world  datasets from the database [https://archive.ics.uci.edu/ml/index.php]:

\textbf{DS-Random-1} - a set of random data generated using the random module of the Python NumPy library. The data has a uniform distribution and contains values on an interval from 0 to 1.

\textbf{DS-Random-2} - a set of random data generated by the scikit-learn library (version 1.0.2) function \texttt{sklearn.datasets.make\_classification}.

\textbf{DS-SkinSegmentation} - a dataset of dimension 3 (RGB values) to distinguish human skin colour from other objects in images \cite{ds-1}.

\textbf{DS-CarEvaluation} - a dataset of dimension 6, representing data from a car valuation model based on the values of 6 features \cite{ds-2}.

Two kinds of studies were carried out:
\begin{itemize}
\item The time required to find K nearest "neighbours" for a test object was measured;
\item The accuracy of recognition of test objects was studied, based on the analysis as to whether their neighbours belong to certain classes.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Test infrastructure}

The computational experiments were performed on the following test infrastructure (Table \ref{tab:t1}).
\begin{table}[h]
	\centering
    \caption{Test infrastructure}\label{tab:t1}
	\begin{tabular}{|l|l|}
		\hline
		CPU & Intel Core i3-7100 (3.9 HHz) \\
		\hline
		RAM & 8 GB \\ 
		\hline
		Operating system & Windows 10 \\
		\hline
		Compiler & Intel(R) oneAPI DPC++/C++ Compiler, Version 2022.0.0 \\
		\hline
	\end{tabular}
\end{table}

@@

On the whole, the experimental results show that the kNN-ME method based on search without removal (procedure A) works on average 2 times faster than the method that involves removal of elements found   (procedure B). Therefore, we will work with the kNN-ME(A) when investigating classification accuracy. In fact, both variants of the kNN-ME method are 5-7 times faster (depending on the dataset) than the kNN-KD method.

@@

The results of performance comparison of the methods on real-world data are qualitatively similar to those obtained by comparison on test data. The kNN-ME method (implementation A) is 1.5 times faster than implementation B, and both implementations are at least 5 times faster than the kNN-KD method.

@@

\begin{table}[h]
	\centering
    \caption{Classes in the CarEvoluation set}\label{tab:t2}
	\begin{tabular}{|c|c|}
		\hline
		Object class & Number of objects in class \\
		\hline
		1 & 1400 \\ 
		\hline
		2 & 400 \\
		\hline
		3 & 70 \\
		\hline
		4 & 65 \\
		\hline
	\end{tabular}
\end{table}

@@

Unfortunately, determining the class of the test object is prone to error. Errors may occur because of the "location" of the object under test on the hypercube edge in immediate proximity to objects of another class. Figure \ref{fig7} shows an example of a random test set where some objects of one class are surrounded by objects of an "alien" class.

\begin{figure}[h]
\widefigure
\includegraphics[width=0.5\linewidth]{fig7.png}
\caption{"Surrounded by aliens"\label{fig7}}
\end{figure}

@@

\begin{itemize}
\item The source dataset was divided into two samples: a training sample and a test sample. The test sample size was 10\% of the initial number of objects.
\item $K$ neighbours were searched for each object in the test sample.
\item The class of an object under test was determined from the set of neighbours found.
\item The recognition result was checked against the known test object class and an error was recorded if there was a discrepancy with the result. The errors were summed up for the whole test sample.
\end{itemize}

@@

\begin{table}[h]
    \caption{Percentage of errors in DS-Random-2 recognition}\label{tab:t3}
	%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
	\begin{tabular} {|p{0,8cm}|p{1,2cm}|p{0,8cm}|p{0,8cm}|p{1,2cm}|p{0,8cm}|p{0,8cm}|p{1,2cm}|p{0,8cm}|p{0,8cm}|}
\hline
K & KDTree (N=2) & A (N=2) & B N=2) & KDTree (N=3) & A (N=3) & B (N=3) & KDTree (N=4) & A (N=4) & B (N=4) \\
\hline
5 & 5 & 5 & 5 & 11 & 12 & 12 & 6 & 8 & 8 \\
\hline
15 & 4 & 6 & 6 & 10 & 11 & 11 & 10 & 9 & 9 \\
\hline
25 & 5 & 7 & 7 & 10 & 11 & 11 & 7 & 9 & 9 \\
\hline
35 & 6 & 6 & 6 & 11 & 13 & 13 & 7 & 10 & 10 \\
\hline
45 & 7 & 6 & 6 & 10 & 13 & 13 & 7 & 10 & 10 \\
\hline
	\end{tabular}
\end{table}

@@

\begin{table}[h]
    \centering
    \caption{Percentage of DS-Skin recognition errors}\label{tab:t4}
	%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
	\begin{tabular} {|p{1cm}|p{1,2cm}|p{1cm}|p{1cm}|}
\hline
K & KDTree (N=3) & A (N=3) & B (N=3) \\
\hline
5 & 0.5 & 0.5 & 0.5 \\
\hline
15 & 0.5 & 0.5 & 0.5 \\
\hline
25 & 0.5 & 0.5 & 0.5 \\
\hline
35 & 1 & 1 & 1 \\
\hline
45 & 1 & 1 & 1 \\
\hline
	\end{tabular}
\end{table}

@@

\begin{table}[h]
    \centering
    \caption{Average error rate (in percent) for DS-CarEvoluation recognition}\label{tab:t5}
	%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
	\begin{tabular} {|p{1cm}|p{1,2cm}|p{1cm}|p{1cm}|}
\hline
K & KDTree (N=6) & A (N=6) & B (N=6) \\
\hline
5 & 8 & 11 & 11 \\
\hline
7 & 9 & 10 & 10 \\
\hline
9 & 8 & 10 & 10 \\
\hline
11 & 10 & 11 & 10 \\
\hline
13 & 9 & 11 & 11 \\
\hline
	\end{tabular}
\end{table}

@@

The paper considers a classical machine learning method, the k nearest neighbour (kNN) method, which is widely used in machine learning practice and has been proven to perform very well because of its "explainability" (the decision to select an object class can be clearly explained). The results of practical implementation of one of kNN acceleration approaches based on dimensionality reduction using space-filling curves are presented.

>>>>

The study was performed using several datasets, both randomly generated for testing and real-world  datasets from the database [https://archive.ics.uci.edu/ml/index.php]:
\begin{enumerate}
\item
\textbf{DS-Random-1} -- a set of random data generated using the random module of the Python NumPy library. The data has a uniform distribution and contains values on an interval from 0 to 1.

\item
\textbf{DS-Random-2} -- a set of random data generated by the scikit-learn library (version 1.0.2) function \texttt{sklearn.datasets.make\_classification}.

\item
\textbf{DS-SkinSegmentation} -- a dataset of dimension 3 (RGB values) to distinguish human skin colour from other objects in images \cite{ds-1}.

\item
\textbf{DS-CarEvaluation} -- a dataset of dimension 6, representing data from a car valuation model based on the values of 6 features \cite{ds-2}.
\end{enumerate}

Two kinds of studies were carried out:
\begin{itemize}
\item The time required to find K nearest neighbours for a test object was measured;
\item The accuracy of recognition of test objects was studied, based on the analysis as to whether their neighbours belong to certain classes.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Test infrastructure}

The computational experiments were performed on the following test infrastructure (Table \ref{tab:t1}).

\begin{specialtable}[H] 
    \caption{Test infrastructure}\label{tab:t1}
	\center
\begin{tabular}{ll}
\toprule
	  Computer specifications & \\
\midrule													
		CPU & Intel Core i3-7100 (3.9 HHz) \\
		RAM & 8 GB \\ 
		Operating system & Windows 10 \\
		Compiler & Intel(R) oneAPI DPC++/C++ Compiler, Version 2022.0.0 \\
\bottomrule
\end{tabular}
\end{specialtable}

@@

On the whole, the experimental results show that the kNN-ME method based on search without removal (procedure A) works on average 2 times faster than the method that involves removal of elements found (procedure B). In fact, both variants of the kNN-ME method are 5-7 times faster (depending on the dataset) than the kNN-KD method.

@@

The results of performance comparison of the methods on real-world data are qualitatively similar to those obtained by comparison on test data. The kNN-ME method (implementation A) is 1.5 times faster than implementation B, and both implementations are at least 5 times faster than the kNN-KD method. Therefore, we will work with the kNN-ME(A) when investigating classification accuracy.

@@

\begin{specialtable}[H] 
  \caption{Classes in the CarEvaluation set}\label{tab:t2}
	\center
\begin{tabular}{cc}
\toprule
	  Object class & Number of objects in class \\
\midrule													
		1 & 1400 \\ 
		2 & 400 \\
		3 & 70 \\
		4 & 65 \\
\bottomrule
\end{tabular}
\end{specialtable}

@@

Unfortunately, determining the class of the test object is prone to error. Errors may occur because of the ``location'' of the object under test on the hypercube edge in immediate proximity to objects of another class. Figure \ref{fig7} shows an example of a random test set where some objects of one class are surrounded by objects of an ``alien'' class.

\begin{figure}[h]
\widefigure
\includegraphics[width=0.7\linewidth]{fig7.png}
\caption{``Surrounded by aliens''\label{fig7}}
\end{figure}

@@

\begin{enumerate}
\item The source dataset was divided into two samples: a training sample and a test sample. The test sample size was 10\% of the initial number of objects.
\item $K$ neighbours were searched for each object in the test sample.
\item The class of an object under test was determined from the set of neighbours found.
\item The recognition result was checked against the known test object class and an error was recorded if there was a discrepancy with the result. The errors were summed up for the whole test sample.
\end{enumerate}

@@

\begin{specialtable}[H] 
  \caption{Percentage of errors in DS-Random-2 recognition}\label{tab:t3}
	\center
	\begin{tabular}{ccccccccc}
\toprule
K & \multicolumn{2}{c}{ $N=2$ } & & \multicolumn{2}{c}{$N=3$} & & \multicolumn{2}{c}{$N=4$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
  & kNN-KD & kNN-ME & & kNN-KD & kNN-ME & & kNN-KD & kNN-ME \\
\midrule													
5  & 5 & 5 & & 11 & 12 & & 6 & 8 \\
15 & 4 & 6 & & 10 & 11 & & 10 & 9 \\
25 & 5 & 7 & & 10 & 11 & & 7 & 9 \\
35 & 6 & 6 & & 11 & 13 & & 7 & 10 \\
45 & 7 & 6 & & 10 & 13 & & 7 & 10 \\
\bottomrule
\end{tabular}
\end{specialtable}

@@

\begin{specialtable}[H] 
  \caption{Percentage of DS-Skin recognition errors, $N=3$}\label{tab:t4}
	\center
\begin{tabular}{ccc}
\toprule
K & kNN-KD & kNN-ME \\
\midrule													
5 & 0.5 & 0.5 \\
15 & 0.5 & 0.5 \\
25 & 0.5 & 0.5\\
35 & 1 & 1\\
45 & 1 & 1\\
\bottomrule
\end{tabular}
\end{specialtable}

@@

\begin{specialtable}[H] 
  \caption{Average error rate (in percent) for DS-CarEvaluation recognition, $N = 6$}\label{tab:t5}
	\center
\begin{tabular}{ccc}
\toprule
K & kNN-KD & kNN-ME \\
\midrule													
5 & 8 & 11 \\
7 & 9 & 10 \\
9 & 8 & 10 \\
11 & 10 & 11 \\
13 & 9 & 11 \\
\bottomrule
\end{tabular}
\end{specialtable}

@@

The paper considers a classical machine learning method, the k nearest neighbour (kNN) method, which is widely used in machine learning practice and has been proven to perform very well because of its ``explainability'' (the decision to select an object class can be clearly explained). The results of practical implementation of one of kNN acceleration approaches based on dimensionality reduction using space-filling curves are presented.
